### 2013-10-07 manuscripts/dpp-msbayes/dpp-msbayes.tex
Fixing error in methods for dpp ms.
Our prior on the concentration parameter of the Dirichlet process was
different between the validation and power analyses, because the
simulated data structure was different (8 pairs vs 22 pairs).

\subsection*{Assessing power}
We evaluated the power of the same four models (Table~\ref{tabPriors}) to
detect random variation in divergence times using methods similar to
\citet{Oaks2012}.
For all of our power simulations, we used a simulated data structure identical
to that of the empirical dataset of Philippine vertebrates analyzed by
\citet{Oaks2012}, which consists of 22 pairs of populations.
For each of the four models, we generated $2\e6$ samples from the prior.
We then generated pseudo-observed datasets from three series of models.
One series of models, which we denote \powerSeriesOld, is identically
distributed as \modelOld except that the divergence times for each of the 22
pairs of populations are randomly drawn from a series of uniform distributions,
$U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and
2.0, in \globalcoalunit generations.
A second series of models, \powerSeriesUniform, are identically distributed as
\modelUniform except that the 22 divergence times are randomly drawn from the
same series of uniform priors as above.
The third series of models, \powerSeriesExp, is also identically
distributed as \modelUniform except that the 22 divergence times are randomly
drawn from a series of of exponential distributions: $Exp(17.3205)$,
$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.3301)$, $Exp(3.4641)$, and $Exp(1.7321)$.
These exponential distributions have the same variance as their uniform
counterparts in the first two series of models.

@@

\begin{table}[htbp]
    \sffamily
    % \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The model priors evaluated in our simulation-based analyses.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
        \midrule
            \modelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,10)$
                      & $\ancestralTheta{} \sim U(0, 0.05)$
                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$
                      & $\divTime{} \sim Exp(0.3464)$
                      & $\ancestralTheta{} \sim Exp(40)$
                      & $\descendantTheta{}{} \sim Exp(40)$ \\
            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
                          & $\divTime{} \sim Exp(0.3464)$
                          & $\ancestralTheta{} \sim Exp(40)$
                          & $\descendantTheta{}{} \sim Exp(40)$ \\
            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
                          & $\divTime{} \sim Exp(0.3464)$
                          & $\ancestralTheta{} \sim Exp(40)$
                          & $\descendantTheta{}{} \sim Exp(40)$ \\
        \bottomrule
    \end{tabular}
    \label{tabPriors}
\end{table}

>>>>

\subsection*{Assessing power}
We evaluated the power of the same four models (Table~\ref{tabPriors}) to
detect random variation in divergence times using methods similar to
\citet{Oaks2012}.
The only difference from the prior models used in the validation analyses above
is the prior on the concentration parameter \concentrationParam for the
\modelDPP model.
Rather than the prior
$\divTimeIndexVector \sim \priorDPP{\sim Gamma(2,2)}$,
we use the prior
$\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
over the discrete divergence models for the model \modelDPP.
The reason for this difference is because for all of our power simulations, we
used a simulated data structure identical to that of the empirical dataset of
Philippine vertebrates analyzed by \citet{Oaks2012}, which consists of 22 pairs
of populations.
For each of the four models, we generated $2\e6$ samples from the prior.
We then generated pseudo-observed datasets from three series of models.
One series of models, which we denote \powerSeriesOld, is identically
distributed as \modelOld except that the divergence times for each of the 22
pairs of populations are randomly drawn from a series of uniform distributions,
$U(0, \divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and
2.0, in \globalcoalunit generations.
A second series of models, \powerSeriesUniform, are identically distributed as
\modelUniform except that the 22 divergence times are randomly drawn from the
same series of uniform priors as above.
The third series of models, \powerSeriesExp, is also identically
distributed as \modelUniform except that the 22 divergence times are randomly
drawn from a series of of exponential distributions: $Exp(17.3205)$,
$Exp(8.6603)$, $Exp(5.7735)$, $Exp(4.3301)$, $Exp(3.4641)$, and $Exp(1.7321)$.
These exponential distributions have the same variance as their uniform
counterparts in the first two series of models.

@@

\begin{table}[htbp]
    \sffamily
    % \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The model priors evaluated in our simulation-based analyses.  For
        the \modelDPP model, the prior on the concentration parameter,
        $\concentrationParam \sim Gamma(\cdot,\cdot)$, was set to $Gamma(2,2)$
        for the validation analyses and $Gamma(1.5,18.1)$ for the power
        analyses.}
    \centering
    \begin{tabular}{ l l l l l }
        \toprule
        & \multicolumn{4}{c}{Priors} \\
        \cmidrule(){2-5}
        Model & \divTimeIndexVector & \divTime{} & \ancestralTheta{} & \descendantTheta{}{} \\
        \midrule
            \modelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,10)$
                      & $\ancestralTheta{} \sim U(0, 0.05)$
                      & $\descendantThetaMean{} \sim U(0, 0.05)$ \\
            \modelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(\cdot,\cdot)}$
                      & $\divTime{} \sim Exp(0.3464)$
                      & $\ancestralTheta{} \sim Exp(40)$
                      & $\descendantTheta{}{} \sim Exp(40)$ \\
            \modelUniform & $\divTimeIndexVector \sim \priorUniform$
                          & $\divTime{} \sim Exp(0.3464)$
                          & $\ancestralTheta{} \sim Exp(40)$
                          & $\descendantTheta{}{} \sim Exp(40)$ \\
            \modelUshaped & $\divTimeIndexVector \sim \priorOld$
                          & $\divTime{} \sim Exp(0.3464)$
                          & $\ancestralTheta{} \sim Exp(40)$
                          & $\descendantTheta{}{} \sim Exp(40)$ \\
        \bottomrule
    \end{tabular}
    \label{tabPriors}
\end{table}
